---
title: "stats220 lab08"
author: "Yuna Otani"
date: "2021/5/16"
output: html_document
---

```{r setup, include = FALSE}
# Setup default locale to en_US in Windows 10
Sys.setlocale(category = "LC_ALL", locale = "English_United States.1252")
library(knitr)
opts_knit$set(root.dir = here::here())
opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.retina = 3,
  comment = "#>")
```

```{r prep}
library(tidyverse)
library(lubridate)
daily_aqi <- read_csv("data/akl-aqi19.csv")
aqi_cat <- fct_inorder(c("Good", "Moderate", "Unhealthy for Sensitive",
                         "Unhealthy", "Very Unhealthy", "Hazardous"))
aqi_pal <- setNames(
  c("#00E400", "#FFFF00", "#FF7E00", "#FF0000", "#8F3F97", "#7E0023"),
  aqi_cat)
```

# Data Preparation
```{r data-prep}
library(dplyr)
get_aqi_cat <- function(x) {                       
  case_when(
    x <= 50 ~ "Good",
    x <= 100 ~ "Moderate",
    x <= 150 ~ "Unhealthy for Sensitive",
    x <= 200 ~ "Unhealthy",
    x <= 300 ~ "Very Unhealthy",
    TRUE ~ "Hazardous"
  ) %>% factor(levels=aqi_cat)
}
daily_aqi <- mutate(daily_aqi,
                    aqi_cat = get_aqi_cat(max_aqi),
                    month = month(date, label=TRUE, abbr=TRUE),
                    mday = mday(date)
)
aqi_table <- group_by(daily_aqi, month) 
aqi_table
```

# QUESTION1
```{r question1}
table_data <- summarise(aqi_table, 
                        Good = sum(aqi_cat == "Good"),
                        Moderate = sum(aqi_cat == "Moderate"),
                        'Unhealthy for Sensitive' = sum(aqi_cat == "Unhealthy for Sensitive"),
                        Unhealthy = sum(aqi_cat == "Unhealthy"),
                        'Very Unhealthy' = sum(aqi_cat == "Very Unhealthy"),
                        Hazardous = sum(aqi_cat == "Hazardous"),
)
kable(table_data) 
```
**In Auckland CBD, on the 23rd and 24th of October a fire broke out on the roof of the New Zealand International Convention Centre. This could be suggested to be one of the reasons why there are 2 days in October where there are very unhealthy aqi conditions.**

# QUESTION2
```{r question2}
daily_aqi %>% 
  ggplot(aes(mday, month, fill = aqi_cat, label = max_aqi)) +
  geom_tile(width = 0.95, height = 0.95) +
  geom_text(size = 3) +
  scale_fill_manual(
    values = aqi_pal, 
    name = 'AQI', 
    drop = FALSE, 
    labels = aqi_cat,
    guide = guide_legend(nrow = 1)
  ) + 
  scale_y_discrete(limits = rev) +
  labs(y = "Month", x = "Day of the Month") +
  coord_fixed(xlim = c(2,30)) +
  theme(
    panel.background = element_rect(fill = NA),
    panel.grid.major = element_line(colour = "grey90"),
    axis.ticks.length.y = unit(0, 'cm'),
    axis.ticks.length.x = unit(0, 'cm'),
    legend.position = 'top'
  )
```

**The bushfires in Australia which occurred during 2019-2020 are believed to be the reason for the extraordinary change in AQI in the second half of 2019.**
